---
title: "Abundance"
author: "Juliana A"
output: html_notebook
---


# Load necessary libraries

```{r}
library(dplyr)
library(tidyr)
library(readr)
```

# Read KEGG data
```{r}
kegg <- read.delim("01_KEGG_Modules", 
                   sep = "\t", 
                   row.names = NULL, 
                   stringsAsFactors = FALSE)
```



# Read Counts data

```{r}
counts <- read.delim("File_D.tsv", sep = "\t", stringsAsFactors = FALSE)
colnames(counts)[1] <- "gene"  # Ensure the first column is named 'gene'
```


# Reshape KEGG data to a long format (one module per row)

```{r}
kegg_long <- kegg %>%
  pivot_longer(cols = starts_with("KEGG_Module"), 
               names_to = "Module_Type", 
               values_to = "KEGG_Module") %>%
  filter(KEGG_Module != "")  # Remove empty rows
```



# Reshape Counts data to a long format for M01 to M30

```{r}
counts_long <- counts %>%
  pivot_longer(cols = starts_with("M"), 
               names_to = "Sample", 
               values_to = "Abundance")
```



# Join the KEGG and Counts data

```{r}
merged_data <- kegg_long %>%
  left_join(counts_long, 
            by = "gene", 
            relationship = "many-to-many")
```



# Group by KEGG_Module and Sample, then sum the abundances

```{r}
result <- merged_data %>%
  group_by(KEGG_Module, Sample) %>%
  summarize(Count = sum(Abundance, na.rm = TRUE), 
            .groups = "drop")
```



# Pivot the result to wide format with KEGG_Module as rows and Samples as columns

```{r}
final_output <- result %>%
  pivot_wider(names_from = Sample, 
              values_from = Count, 
              values_fill = list(Count = 0))
```



# Print the final output

```{r}
print(final_output)
```


# Optionally write to a file

```{r}
write.table(final_output, "KEGG_Counts_Output.txt", sep = "\t", row.names = FALSE, quote = FALSE)

```

